---
title: "Notes - Douglas Luke, A User's Guide to Network Analysis in R"
author: "Mustafa Ascha"
date: "August 21, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(statnet)
library(UserNetR)
```

## Chapter 2

Jacob Moreno collected data on friendship networks of fourth graders (in the 1930s lol). Here's a visualization, according to gender:

```{r}

data(Moreno)

gender <- Moreno %v% "gender"

plot(Moreno, vertex.col = gender + 2, vertex.cex = 1.2)

```

Okay, so that's hilarious. It looks like there was only one pair of boy-girl friends. 

## Describing the previous network

Use the following: 
* Size (number of nodes)
* Density (Proportion of potential connections)
* Components (Number of distinct sub-networks)
* Diameter (Compactness)
* Clustering Coefficient

```{r}

summary(Moreno, print.adj = FALSE)

```

`r network.size(Moreno)` is the network **size** of Moreno. The previous summary shows some of the other attributes. 

Note that the maximum number of edges among nodes in a **directed** graph is *k* `*` *(k - 1)*. Where *L* is the number of observed edges, the **density** is *L* / *k* `*` *(k - 1)*. 

Similarly the maximum number of edges among nodes in an **undirected** graph is *(k* `*` *(k - 1)) / 2*. Where *L* is the number of observed edges, the **density** is *2L* / *k* `*` *(k - 1)*.

`r gden(Moreno)` is the network **density** of Moreno. 

A component is a type of subgroup, though there are many other ways to define a subgroup. A component is defined as a group of nodes that has no connections to any other nodes. There are `r components(Moreno)` **components** in Moreno. 

Diameter is another interesting measure. Of the shortest paths across all pairs of nodes, the longest is called the diameter. This is a worst-case scenario for the transferral of information, so it kinda represents the farthest a message could travel if everyone wanted to send messages to everyone else. A shortest path is called a "geodesic", so the diameter is the maximum geodesic. The **diameter** of Moreno is `r max(geodist(component.largest(Moreno, result = "graph"))$gdist)`. 

To understand the clustering coefficient, first consider closure: closure is the process whereby two people who share a common friend become friends, themselves. This is measured in terms of *transitivity*, which is the proportion of closed three-person relationships to the proportion of all three-person relationships. The **transitivity** of Moreno is `r gtrans(Moreno, mode = "graph")`. This is one type of clustering coefficient. 

## Chapter 3

There are two text representations of networks to know well--adjacency matrices and edge lists. 

An adjacency matrix is a square matrix with a 1 representing each connection (for directed graphs, going from row -> column). Undirected graphs have triangular matrices. 

An edge list is a two-variable table where one variable represents a starting node, and the second variable represents the end node.

Network formats include Pajek (.net or .paj files), GraphML, and UCINet DL files. 

Different packages have different network objects with different attributes and different methods (different redundancy, too...). However, the two packages to know at this point are `statnet` and `igraph`, which can work together using the `intergraph` package. 

Chapter 3 seems mainly useful insofar as it introduces various network object attributes. 

* `network.vertex.names(networkObject)` pulls the vertex names
* `as.sociomatrix(networkObject)` produces an adjacency matrix--this reminds me of the reshaping functions in tidyr and reshape2
* `set.vertex.attribute(networkObject, attribute, attributeValues)` would set the attribute of networkObject vertices to attributeValues. Rather than setting attributes, I should look into how I can read edge list data with attributes as extra columns
* `get.vertex.attribute(networkObject, attribute)` and `networkObject %v% attribute` do the same thing, but `%v%` is an infix operator
* `list.vertex.attributes(networkObject)` shows the verex attributes present in a network object. 
* `list.edge.attributes(networkObject)`, `get.edge.attributes(networkObject, attribute)`, `networkObject %e% attribute`, and `set.edge.attribute(networkObject, attribute, attributeValues)` are the corresponding edge functions 

Edge attributes come in handy when working with **valued networks**, which have some numeric value associated with the edge. This is interesting/useful. 

For an adjacency matrix with valued edges (instead of the binary values presented earlier), the network can be read by:

```{r valuedRead}
netVal <- rbind(c(0,2,3,0,0),
                c(0,0,3,1,0),
                c(0,1,0,0,0),
                c(0,0,0,0,0),
                c(0,0,2,0,0))

netval <- network(netval, matrix.type = "adjacency", 
                  ignore.eval = FALSE, names.eval = "name")
```

And now, netVal edges have a "name" attribute. 

Here's an example using the `statnet` package:

```{r statnetChapter3}

library(statnet)



```







