---
title: "Notes - Douglas Luke, A User's Guide to Network Analysis in R"
author: "Mustafa Ascha"
date: "August 21, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(statnet)
library(UserNetR)
```

## Chapter 2

Jacob Moreno collected data on friendship networks of fourth graders (in the 1930s lol). Here's a visualization, according to gender:

```{r}

data(Moreno)

gender <- Moreno %v% "gender"

plot(Moreno, vertex.col = gender + 2, vertex.cex = 1.2)

```

Okay, so that's hilarious. It looks like there was only one pair of boy-girl friends. 

## Describing the previous network

Use the following: 
* Size (number of nodes)
* Density (Proportion of potential connections)
* Components (Number of distinct sub-networks)
* Diameter (Compactness)
* Clustering Coefficient

```{r}

summary(Moreno, print.adj = FALSE)

```

`r network.size(Moreno)` is the network **size** of Moreno. The previous summary shows some of the other attributes. 

Note that the maximum number of edges among nodes in a **directed** graph is *k* `*` *(k - 1)*. Where *L* is the number of observed edges, the **density** is *L* / *k* `*` *(k - 1)*. 

Similarly the maximum number of edges among nodes in an **undirected** graph is *(k* `*` *(k - 1)) / 2*. Where *L* is the number of observed edges, the **density** is *2L* / *k* `*` *(k - 1)*.

`r gden(Moreno)` is the network **density** of Moreno. 

A component is a type of subgroup, though there are many other ways to define a subgroup. A component is defined as a group of nodes that has no connections to any other nodes. There are `r components(Moreno)` **components** in Moreno. 

Diameter is another interesting measure. Of the shortest paths across all pairs of nodes, the longest is called the diameter. This is a worst-case scenario for the transferral of information, so it kinda represents the farthest a message could travel if everyone wanted to send messages to everyone else. A shortest path is called a "geodesic", so the diameter is the maximum geodesic. The **diameter** of Moreno is `r max(geodist(component.largest(Moreno, result = "graph"))$gdist)`. 

To understand the clustering coefficient, first consider closure: closure is the process whereby two people who share a common friend become friends, themselves. This is measured in terms of *transitivity*, which is the proportion of closed three-person relationships to the proportion of all three-person relationships. The **transitivity** of Moreno is `r gtrans(Moreno, mode = "graph")`. This is one type of clustering coefficient. 

## Chapter 3









